# --------------------------------------------
# ---------- 12_DataVisualization_ggplot2_part_1 (dataviz) ----
# --------------------------------------------
# Principles;
	# Graphics = distinct layers of grammatical elements
	# Meaningful plots through aesthetic mapping

# Essential     Grammatical Elements
# Element			  Description
# Data			    The dataset being plotted
# Aesthetics		The scales onto which we map our dataset
# Geometries		The visual elements used for our dataset
# Other elements
# Facets		    Plotting small multiples
# Statistics		Presentations of our data to aid understanding
# Coordinates		The space on which the data will be plotted
# Themes			  All non-data ink

# -------------- Introduction ----------------------
```{r setup}
library(ggplot2); library(tidyr); library(tidyverse)
```


```{r}
# (miles per gallon) against the weight (in thousands of pounds)
# A scatter plot has been made for you
# aes = aesthetics
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point()
# Replace ___ with the correct column
ggplot(mtcars, aes(x = wt, y = mpg, color = disp)) +
  geom_point()
# Replace ___ with the correct column
ggplot(mtcars, aes(x = wt, y = mpg, size = disp)) +
  geom_point()
```


# A continuous variable can not be mapped to shape
# data sets used
# https://warwick.ac.uk/fac/sci/moac/people/students/peter_cock/r/iris_plots/
```{r}
head(iris); summarize(iris) # in R
library(ggplot2); library('ggplot2')
head(diamonds) # in R
```

# Explore the diamonds data frame with str()
str(diamonds)
# Add geom_point() with +
```{r}
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point() 
# Add geom_point() and geom_smooth() with +
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point() +
  geom_smooth()
```

```{r}
# 1 - The plot you created in the previous exercise
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point() +
  geom_smooth()
# 2 - Copy the above command but show only the smooth line
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_smooth()
# 3 - Copy the above command and assign the correct value to col in aes()
ggplot(diamonds, aes(x = carat, y = price, color = clarity)) +
  geom_smooth()
# 4 - Keep the color settings from previous command. Plot only the points with argument alpha.
ggplot(diamonds, aes(x = carat, y = price, color = clarity)) +
  geom_point(alpha = 0.4) 
```

```{r}
# Create the object containing the data and aes layers: dia_plot
dia_plot <- ggplot(diamonds, aes(x = carat, y = price))
# Add a geom layer with + and geom_point()
dia_plot + geom_point()
# Add the same geom layer, but with aes() inside
dia_plot + geom_point(aes(color = clarity))
```

```{r}
# 1 - The dia_plot object has been created for you
dia_plot <- ggplot(diamonds, aes(x = carat, y = price))
# 2 - Expand dia_plot by adding geom_point() with alpha set to 0.2
dia_plot <- dia_plot + geom_point(alpha = 0.2)
# 3 - Plot dia_plot with additional geom_smooth() with se set to FALSE
dia_plot + geom_smooth(se = FALSE)
# 4 - Copy the command from above and add aes() with the correct mapping to geom_smooth()
dia_plot + geom_smooth(aes(col = clarity), se = FALSE)
```

```{r}
# ---- Data -----------------
# base plot
# Plot the correct variables of mtcars
plot(mtcars$wt, mtcars$mpg, col = mtcars$cyl)
# Change cyl inside mtcars to a factor
mtcars$fcyl <- as.factor(mtcars$cyl)
# Make the same plot as in the first instruction
plot(mtcars$wt, mtcars$mpg, col = mtcars$fcyl)
```

```{r}
# Use lm() to calculate a linear model and save it as carModel
carModel <- lm(mpg ~ wt, data = mtcars)
# Basic plot
mtcars$cyl <- as.factor(mtcars$cyl)
plot(mtcars$wt, mtcars$mpg, col = mtcars$cyl)
# Call abline() with carModel as first argument and set lty to 2
abline(carModel, lty = 2)
# Plot each subset efficiently with lapply
# You don't have to edit this code
plot(mtcars$wt, mtcars$mpg, col = mtcars$cyl)
lapply(mtcars$cyl, function(x) {
  abline(lm(mpg ~ wt, mtcars, subset = (cyl == x)), col = x)
})
# This code will draw the legend of the plot
# You don't have to edit this code
legend(x = 5, y = 33, legend = levels(mtcars$cyl),
       col = 1:3, pch = 1, bty = "n")
```


```{r}
# Convert cyl to factor (don't need to change)
mtcars$cyl <- as.factor(mtcars$cyl)
# Example from base R (don't need to change)
plot(mtcars$wt, mtcars$mpg, col = mtcars$cyl)
abline(lm(mpg ~ wt, data = mtcars), lty = 2)
lapply(mtcars$cyl, function(x) {
  abline(lm(mpg ~ wt, mtcars, subset = (cyl == x)), col = x)
})
legend(x = 5, y = 33, legend = levels(mtcars$cyl),
       col = 1:3, pch = 1, bty = "n")
```

```{r}
# Plot 1: add geom_point() to this command to create a scatter plot
ggplot(mtcars, aes(x = wt, y = mpg, col = cyl)) +
  geom_point()
# Plot 2: include the lines of the linear models, per cyl
ggplot(mtcars, aes(x = wt, y = mpg, col = cyl)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
# Plot 3: include a lm for the entire dataset in its whole
ggplot(mtcars, aes(x = wt, y = mpg, col = cyl)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE, linetype = 2)
```


```{r}
# --  tidy data()
# Consider the structure of iris, iris.wide and iris.tidy (in that order)
str(iris)
str(iris.wide)
str(iris.tidy)

# Think about which dataset you would use to get the plot shown right
# Fill in the ___ to produce the plot given to the right
ggplot(iris.tidy, aes(x = Species, y = Value, col = Part)) +
  geom_jitter() +
  facet_grid(. ~ Measure)
```

```{r}
# Load the tidyr package
library(tidyr)
# The 3 data frames (iris, iris.wide and iris.tidy) are available in your environment
# Execute head() on iris, iris.wide and iris.tidy (in that order)
head(iris)
head(iris.wide) # <--------
head(iris.tidy)
```

```{r}
# Think about which dataset you would use to get the plot shown right
# Fill in the ___ to produce the plot given to the right
ggplot(iris.wide, aes(x = Length, y = Width, color = Part)) +
  geom_jitter() +
  facet_grid(. ~ Species)

```

```{r}
# Load the tidyr package
library(tidyr)

# Add column with unique ids (don't need to change)
iris$Flower <- 1:nrow(iris)

# Fill in the ___ to produce to the correct iris.wide dataset
iris.wide <- iris %>%
  gather(key, value, -Species, -Flower) %>%
  separate(key, c("Part", "Measure"), "\\.") %>%
  spread(Measure, value)
```

```{r}
# ---- Aesthetics -----------------
# 1 - Map mpg to x and cyl to y
ggplot(mtcars, aes(mpg, cyl)) +
  geom_point()
# 2 - Reverse: Map cyl to x and mpg to y
ggplot(mtcars, aes(cyl, mpg)) +
  geom_point()
# 3 - Map wt to x, mpg to y and cyl to col
ggplot(mtcars, aes(wt, mpg, color = cyl)) +
  geom_point()
# 4 - Change shape and size of the points in the above plot
ggplot(mtcars, aes(wt, mpg, color = cyl)) +
  geom_point(shape = 1, size = 4)
```

```{r}
# am and cyl are factors, wt is numeric
class(mtcars$am)
class(mtcars$cyl)
class(mtcars$wt)
```

```{r}
# From the previous exercise
ggplot(mtcars, aes(x = wt, y = mpg, col = cyl)) +
  geom_point(shape = 1, size = 4)
# 1 - Map cyl to fill
ggplot(mtcars, aes(x = wt, y = mpg, fill = cyl)) +
  geom_point(shape = 1, size = 4)
# 2 - Change shape and alpha of the points in the above plot
ggplot(mtcars, aes(x = wt, y = mpg, fill = cyl)) +
  geom_point(shape = 21, size = 4, alpha = 0.6)
# 3 - Map am to col in the above plot
ggplot(mtcars, aes(x = wt, y = mpg, fill = cyl, col = am)) +
  geom_point(shape = 21, size = 4, alpha = 0.6)
```

```{r}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point(size = cyl)
# Map cyl to size
ggplot(mtcars, aes(x = wt, y = mpg, size = cyl)) +
  geom_point()
# Map cyl to alpha
ggplot(mtcars, aes(x = wt, y = mpg, alpha = cyl)) +
  geom_point()
# Map cyl to shape 
ggplot(mtcars, aes(x = wt, y = mpg, shape = cyl)) +
  geom_point()
# Map cyl to label
ggplot(mtcars, aes(x = wt, y = mpg, label = cyl)) +
  geom_point() +
  geom_text()
```

```{r}
# Define a hexadecimal color
# install.packages("colourpicker")
# library(colourpicker)
# colourpicker() # plugin returns = c("#EEDFCC", "#7FFF00", "#8B1A1A", "#F0FFFF")

my_color <- "#4ABEFF"

# Draw a scatter plot with color *aesthetic*
ggplot(mtcars, aes(x = wt, y = mpg, size = cyl, col = cyl)) +
  geom_point()
# Same, but set color *attribute* in geom layer 
ggplot(mtcars, aes(x = wt, y = mpg, size = cyl)) +
  geom_point(col = my_color)
# Set the fill aesthetic; color, size and shape attributes
ggplot(mtcars, aes(x = wt, y = mpg, fill = cyl)) +
  geom_point(col = my_color, size = 10, shape = 23)
# Expand to draw points with alpha 0.5
ggplot(mtcars, aes(x = wt, y = mpg, fill = cyl)) +
  geom_point( alpha = 0.5)
# Expand to draw points with shape 24 and color yellow
ggplot(mtcars, aes(x = wt, y = mpg, fill = cyl)) +
  geom_point(shape = 24, col = "yellow")
# Expand to draw text with label rownames(mtcars) and color red
ggplot(mtcars, aes(x = wt, y = mpg, fill = cyl, rownames(mtcars), col = "red")) +
  geom_point()
```

```{r}
# Map mpg onto x, qsec onto y and factor(cyl) onto col
ggplot(mtcars, aes(x = mpg, y = qsec, col = factor(cyl))) +
  geom_point()
# Add mapping: factor(am) onto shape
ggplot(mtcars, aes(x = mpg, y = qsec, col = factor(cyl), shape = factor(am))) +
  geom_point()
# Add mapping: (hp/wt) onto size
ggplot(mtcars, aes(x = mpg, y = qsec, col = factor(cyl), shape = factor(am), size = (hp/wt))) +
  geom_point()
```

```{r}
# 1 - Create jittered plot of mtcars, mpg onto x, 0 onto y
ggplot(mtcars, aes(x = mpg, y = 0)) +
  geom_jitter()
# 2 - Add function to change y axis limits
ggplot(mtcars, aes(x = mpg, y = 0)) +
  geom_jitter() +
  scale_y_continuous(limits = c(-2,2))
```


```{r}
# Basic scatter plot: wt on x-axis and mpg on y-axis; map cyl to col
ggplot(mtcars, aes(x = wt, y = mpg, col = cyl)) +
  geom_point(size=4)
# Hollow circles - an improvement
ggplot(mtcars, aes(x = wt, y = mpg, col = cyl)) +
  geom_point(size=4,shape = 1)
# Add transparency - very nice
ggplot(mtcars, aes(x = wt, y = mpg, col = cyl)) +
  geom_point(size=4,alpha = 0.6)
```

```{r}
# Scatter plot: carat (x), price (y), clarity (color)
ggplot(diamonds, aes(x = carat, y = price, col = clarity)) +
  geom_point()
# Adjust for overplotting
ggplot(diamonds, aes(x = carat, y = price, col = clarity)) +
  geom_point(alpha = 0.5)
# Scatter plot: clarity (x), carat (y), price (color)
ggplot(diamonds, aes(x = clarity, y = carat, col = price)) +
  geom_point(alpha = 0.5)
# Dot plot with jittering
ggplot(diamonds, aes(x = clarity, y = carat, col = price)) +
  geom_point(alpha = 0.5, position = "jitter")
```

```{r}
# ---- Geometries -----------------
# Shown in the viewer:
ggplot(mtcars, aes(x = cyl, y = wt)) +
  geom_point()
# Solutions:
# 1 - With geom_jitter()
ggplot(mtcars, aes(x = cyl, y = wt)) +
  geom_jitter()
# 2 - Set width in geom_jitter()
ggplot(mtcars, aes(x = cyl, y = wt)) +
  geom_jitter(width = 0.1)
# 3 - Set position = position_jitter() in geom_point()
ggplot(mtcars, aes(x = cyl, y = wt)) +
  geom_point(position = position_jitter(0.1))
```

```{r}
# Examine the structure of Vocab
str(Vocab)
# Basic scatter plot of vocabulary (y) against education (x). Use geom_point()
ggplot(Vocab, aes(x = education, y = vocabulary)) +
  geom_point()
# Use geom_jitter() instead of geom_point()
ggplot(Vocab, aes(x = education, y = vocabulary)) +
  geom_jitter()
# Using the above plotting command, set alpha to a very low 0.2
ggplot(Vocab, aes(x = education, y = vocabulary)) +
  geom_jitter(alpha = 0.2)
# Using the above plotting command, set the shape to 1
ggplot(Vocab, aes(x = education, y = vocabulary)) +
  geom_jitter(alpha = 0.2, shape = 1)
```

```{r}
# 1 - Make a univariate histogram
ggplot(mtcars, aes(x = mpg)) +
  geom_histogram()
# 2 - Plot 1, plus set binwidth to 1 in the geom layer
ggplot(mtcars, aes(x = mpg)) +
  geom_histogram(binwidth = 1)
# 3 - Plot 2, plus MAP ..density.. to the y aesthetic (i.e. in a second aes() function)
ggplot(mtcars, aes(x = mpg)) +
  geom_histogram(aes( y = ..density..),binwidth = 1)
# 4 - plot 3, plus SET the fill attribute to "#377EB8"
ggplot(mtcars, aes(x = mpg)) +
  geom_histogram(aes( y = ..density..),binwidth = 1, fill="#377EB8" )
```

```{r}
# Three position arguments will be introduced here:
#   
#   stack: place the bars on top of each other. Counts are used. This is the default position.
# fill: place the bars on top of each other, but this time use proportions.
# dodge: place the bars next to each other. Counts are used.

# Draw a bar plot of cyl, filled according to am
ggplot(mtcars, aes(x = cyl, fill = am)) +
  geom_bar()
# Change the position argument to stack
ggplot(mtcars, aes(x = cyl, fill = am)) +
  geom_bar(position = "stack")
# Change the position argument to fill
ggplot(mtcars, aes(x = cyl, fill = am)) +
  geom_bar(position = "fill")
# Change the position argument to dodge
ggplot(mtcars, aes(x = cyl, fill = am)) +
  geom_bar(position = "dodge")
```

```{r}
# Remember, the reason you want to use position_dodge() 
# (and position_jitter()) is to specify how much dodging (or jittering) you want.
# 1 - The last plot form the previous exercise
ggplot(mtcars, aes(x = cyl, fill = am)) +
  geom_bar(position = "dodge")
# 2 - Define posn_d with position_dodge()
posn_d <- position_dodge(width = 0.2)
# 3 - Change the position argument to posn_d
ggplot(mtcars, aes(x = cyl, fill = am)) +
  geom_bar(position = posn_d)
# 4 - Use posn_d as position and adjust alpha to 0.6
ggplot(mtcars, aes(x = cyl, fill = am)) +
  geom_bar(position = posn_d, alpha = 0.6)
```

```{r}


# Overlapping histograms pose similar problems to overlapping bar plots, 
# but there is a unique solution here: a frequency polygon.
# This is a geom specific to binned data that draws a line connecting the value of each bin. 
#   Like geom_histogram(), it takes a binwidth argument and by default stat = "bin" and position = "identity".
# A basic histogram, add coloring defined by cyl
ggplot(mtcars, aes(mpg, fill = cyl)) +
  geom_histogram(binwidth = 1)
# Change position to identity
ggplot(mtcars, aes(mpg, fill = cyl)) +
  geom_histogram(binwidth = 1,position = "identity")
# Change geom to freqpoly (position is identity by default)
ggplot(mtcars, aes(mpg, color = cyl)) +
  geom_histogram(binwidth = 1,position = "identity") +
  geom_freqpoly(binwidth = 1)
```

```{r}
# Bar plots with color ramp, part 1
# 
# In this example of a bar plot, you'll fill each segment 
#   according to an ordinal variable. The best way 
#   to do that is with a sequential color series.
# You'll be using the Vocab dataset from earlier. Since this is a 
#   much larger dataset with more categories, 
#   you'll also compare it to a simpler dataset, mtcars. 
#   Both datasets are ordinal.
# Example of how to use a brewed color palette
ggplot(mtcars, aes(x = cyl, fill = am)) +
  geom_bar() +
  scale_fill_brewer(palette = "Set1")
# Use str() on Vocab to check out the structure
str(Vocab)
# Plot education on x and vocabulary on fill
# Use the default brewed color palette
ggplot(Vocab,aes(x = education, fill = vocabulary)) +
  geom_bar(position = "fill") +
  scale_fill_brewer()  
# the plot is not complete. Let's fix this in the next exercise. 
# This was because for continuous data, the default RColorBrewer palette 
#   that scale_fill_brewer() calls is "Blues". There are only 9 colours in the palette, 
#   and since you have 11 categories, your plot looked strange.
new_col <- colorRampPalette(c("#FFFFFF", "#0000FF"))
new_col(21) # the newly extrapolated colours
munsell::plot_hex(new_col(4)) # Quick and dirty plot
```

```{r}
# Final plot of last exercise
ggplot(Vocab, aes(x = education, fill = vocabulary)) +
  geom_bar(position = "fill") +
  scale_fill_brewer()
# Definition of a set of blue colors
blues <- brewer.pal(9, "Blues") # from the RColorBrewer package
# 1 - Make a color range using colorRampPalette() and the set of blues
blue_range <- colorRampPalette(blues)
# 2 - Use blue_range to adjust the color of the bars, use scale_fill_manual()
ggplot(Vocab, aes(x = education, fill = vocabulary)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = blue_range(11))
```

```{r}
# Overlapping histograms (2)
# 1 - Basic histogram plot command
ggplot(mtcars, aes(mpg)) +
  geom_histogram(binwidth = 1)
# 2 - Plot 1, Expand aesthetics: am onto fill
ggplot(mtcars, aes(mpg, fill = am)) +
  geom_histogram(binwidth = 1)
# 3 - Plot 2, change position = "dodge"
ggplot(mtcars, aes(mpg, fill = am)) +
  geom_histogram(binwidth = 1, position = "dodge")
# 4 - Plot 3, change position = "fill"
ggplot(mtcars, aes(mpg, fill = am)) +
  geom_histogram(binwidth = 1, position = "fill")
# 5 - Plot 4, plus change position = "identity" and alpha = 0.4
ggplot(mtcars, aes(mpg, fill = am)) +
  geom_histogram(binwidth = 1, position = "identity", alpha = 0.4)
# 6 - Plot 5, plus change mapping: cyl onto fill
ggplot(mtcars, aes(mpg, fill = cyl)) +
  geom_histogram(binwidth = 1, position = "identity", alpha = 0.4)
```

```{r}
# To explore this topic, you'll use the economics data frame, 
# which contains time series for unemployment and population statistics 
# from the Federal Reserve Bank of St. Louis in the US. 
# The data is contained in the ggplot2 package.
# Print out head of economics
head(economics)
# Plot unemploy as a function of date using a line plot
ggplot(economics, aes(x = date, y = unemploy)) +
  geom_line()
# Adjust plot to represent the fraction of total population that is unemployed
ggplot(economics, aes(x = date, y = unemploy/pop)) +
  geom_line()
```

```{r}
# Basic line plot
ggplot(economics, aes(x = date, y = unemploy/pop)) +
  geom_line()
# Expand the following command with geom_rect() to draw the recess periods
ggplot(economics, aes(x = date, y = unemploy/pop)) +
  geom_rect(data = recess,
            aes(xmin = begin, xmax = end, ymin = -Inf, ymax = +Inf),
            inherit.aes = FALSE, fill = "red", alpha = 0.2) +
  geom_line()
```

```{r}
# Multiple time series, part 1
# Check the structure as a starting point
str(fish.species)
# Use gather to go from fish.species to fish.tidy
fish.tidy <- gather(fish.species, Species, Capture, -Year)
head(fish.tidy)

# Recreate the plot shown on the right
ggplot(fish.tidy, aes(x = Year, y = Capture, color = Species)) +
  geom_line()
```

```{r}
# ---- qplot and wrap-up  -----------------
# ggplot2 offers a powerful and diverse array of functions, 
#   but qplot() allows for quick and dirty plots. 
#   Plus, you should also be familiar with basic plotting notation.
# The old way (shown)
plot(mpg ~ wt, data = mtcars) # formula notation
with(mtcars, plot(wt, mpg)) # x, y notation
# Using ggplot:
ggplot(mtcars, aes(wt, mpg)) +
  geom_point()
# Using qplot:
qplot(wt, mpg, data = mtcars)
```

```{r}
# basic qplot scatter plot:
qplot(wt, mpg, data = mtcars)
# Categorical variable mapped onto size:
# cyl
qplot(wt, mpg, data = mtcars, size = factor(cyl))
# gear
qplot(wt, mpg, data = mtcars, size = factor(gear))
# Continuous variable mapped onto col:
# hp
qplot(wt, mpg, data = mtcars, col = hp)
# qsec
qplot(wt, mpg, data = mtcars, col = qsec)
```

```{r}
# qplot() with x only
qplot(factor(cyl),data = mtcars)
# qplot() with x and y
qplot(factor(cyl),factor(vs),data = mtcars)
# qplot() with geom set to jitter manually
qplot(factor(cyl),factor(vs),data = mtcars, geom = "jitter") 
```

```{r}
# Choosing geoms, part 2 - dotplot
# Some naming conventions:
#   Scatter plots:
#   Continuous x, continuous y.
#   Dot plots:
#   Categorical x, continuous y.

# cyl and am are factors, wt is numeric
class(mtcars$cyl)
class(mtcars$am)
class(mtcars$wt)

# "Basic" dot plot, with geom_point():
ggplot(mtcars, aes(cyl, wt, col = am)) +
  geom_point(position = position_jitter(0.2, 0))
# 1 - "True" dot plot, with geom_dotplot():
ggplot(mtcars, aes(cyl, wt, fill = am)) +
  geom_dotplot(binaxis = "y", stackdir = "center")
# 2 - qplot with geom "dotplot", binaxis = "y" and stackdir = "center"
qplot(
  cyl, wt,
  data = mtcars,
  fill = am,
  geom = "dotplot",
  binaxis = "y",
  stackdir = "center"
)
```

```{r}
# Chicken weight
# 
# The ChickWeight dataset is a data frame which represents 
# the progression of weight of several chicks. 
# The little chicklings are each given a specific diet. 
# There are four types of diet and the farmer wants to know 
# which one fattens the chicks the fastest.
# > head(ChickWeight)
# weight Time Chick Diet
# 1     42    0     1    1
# 2     51    2     1    1
# 3     59    4     1    1
# 4     64    6     1    1
# 5     76    8     1    1
# 6     93   10     1    1
# > 
# ChickWeight is available in your workspace
# 1 - Check out the head of ChickWeight
head(ChickWeight)
# 2 - Basic line plot
ggplot(ChickWeight, aes(x = Time, y = weight)) +
  geom_line(aes(group = Chick))
# 3 - Take plot 2, map Diet onto col.
ggplot(ChickWeight, aes(x = Time, y = weight, color = Diet)) +
  geom_line(aes(group = Chick))
# 4 - Take plot 3, add geom_smooth()
ggplot(ChickWeight, aes(x = Time, y = weight, color = Diet)) +
  geom_line(aes(group = Chick),alpha = 0.3) +
  geom_smooth(lwd = 2, se = FALSE)
```


```{r}
# Titanic
# You've watched the movie Titanic by James Cameron (1997) again and 
#   after a good portion of sobbing you decide to investigate 
#   whether you'd have a chance of surviving this disaster.
# > str(titanic)
# 'data.frame':	714 obs. of  4 variables:
#   $ Survived: int  0 1 1 1 0 0 0 1 1 1 ...
# $ Pclass  : int  3 1 3 1 3 1 3 3 2 3 ...
# $ Sex     : Factor w/ 2 levels "female","male": 2 1 1 1 2 2 2 1 1 1 ...
# $ Age     : num  22 38 26 35 35 54 2 27 14 4 ...
# titanic is avaliable in your workspace
# 1 - Check the structure of titanic
str(titanic)
# 2 - Use ggplot() for the first instruction
ggplot(titanic, aes(x = Pclass, fill = Sex)) +
  geom_bar(position = "dodge")
# 3 - Plot 2, add facet_grid() layer
ggplot(titanic, aes(x = Pclass, fill = Sex)) +
  geom_bar(position = "dodge") +
  facet_grid(. ~ Survived)
# 4 - Define an object for position jitterdodge, to use below
posn.jd <- position_jitterdodge(0.5, 0, 0.6)
# 5 - Plot 3, but use the position object from instruction 4
ggplot(titanic, aes(x = Pclass, y= Age, color = Sex)) +
  geom_point(size = 3, alpha = 0.5, position = posn.jd) +
  facet_grid(. ~ Survived)
```














  
  
  
  
  
  
  
  
